On the uniqueness and stability of dictionaries for sparse representation of noisy signals

### 2017-08-18 manuscripts/PNAS/chazthm_pnas.tex

We furthermore note that unlike in previous works \cite{li2004analysis, Georgiev05, Aharon06, Hillar15}, the matrix $\mathbf{A}$ need not satisfy the spark condition to be recoverable from data. As a simple example for $k=2$, let $\mathbf{A} = [ \mathbf{e}_1, \ldots, \mathbf{e}_5, \mathbf{v}]$ where $\mathbf{v} = \mathbf{e}_1 + \mathbf{e}_3 + \mathbf{e}_5$, and take $\mathcal{H}$ to be all consecutive pairs of $[m]$ arranged in cyclic order. Then, $\mathbf{A}$ satisfies the assumptions of Thm.~\ref{DeterministicUniquenessTheorem} but not \eqref{SparkCondition}. %(since $\{ \mathbf{A}_1, \mathbf{A}_3, \mathbf{A}_5, \mathbf{A}_6\}$ is not a linearly independent set).
%This weakening allows for a practical (polynomial) amount of data to still guarantee a stable dictionary in the case where the set of sparse supports for the $\mathbf{\bar x}_i$ is known to have a size that grows polynomially in $m$ and $k$.

@@

\begin{proposition}\label{neuroprop}
Sparse neural population activity is recoverable from noisy random linear compression by any method that solves Prob.~\ref{DeterministicUniquenessTheorem} or \ref{SLCopt}; in particular, via biophysically plausible unsupervised sparse linear coding (e.g. \cite{rehnsommer2007, rozell2007neurally, ganguli2012compressed}).
\end{proposition}
%\cite{rehnsommer2007, rozell2007neurally, ganguli2012compressed, hu2014hebbian}

We close this section with a comment on sample complexity. A basic open problem remains: how many samples are necessary to determine the sparse linear coding model? If $k$ is held fixed or if the size of the support set of reconstructing codes is known to be polynomial in $\bar m$ and $k$, then a practical (polynomial) amount of data suffices.\footnote{In the latter case, a reexamination of the pigeonholing argument in the proof of Thm.~\ref{DeterministicUniquenessTheorem} requires a polynomial number of samples distributed over a polynomial number of supports.} Reasons to be skeptical that this holds in general, however, can be found in \cite{tillmann2014computational, tillmann2015computational}.

@@

In this section, we prove Lem.~\ref{MainLemma} after stating auxiliary lemmas.% and then sketch the proofs of Thm.~\ref{robustPolythm} and Cor.~\ref{ProbabilisticCor}.

>>>>

We furthermore note that unlike in previous works \cite{li2004analysis, Georgiev05, Aharon06, Hillar15}, the matrix $\mathbf{A}$ need not satisfy the spark condition to be recoverable from data. As a simple example for $k=2$, let $\mathbf{A} = [ \mathbf{e}_1, \ldots, \mathbf{e}_5, \mathbf{v}]$ where $\mathbf{v} = \mathbf{e}_1 + \mathbf{e}_3 + \mathbf{e}_5$, and take $\mathcal{H}$ to be all consecutive pairs of $[m]$ arranged in cyclic order. Then, dictionary $\mathbf{A}$ satisfies the assumptions of Thm.~\ref{DeterministicUniquenessTheorem} but not \eqref{SparkCondition}. %(since $\{ \mathbf{A}_1, \mathbf{A}_3, \mathbf{A}_5, \mathbf{A}_6\}$ is not a linearly independent set).
%This weakening allows for a practical (polynomial) amount of data to still guarantee a stable dictionary in the case where the set of sparse supports for the $\mathbf{\bar x}_i$ is known to have a size that grows polynomially in $m$ and $k$.

@@

\begin{proposition}\label{neuroprop}
Sparse neural population activity is recoverable from noisy random linear compression by any method that solves Prob.~\ref{DeterministicUniquenessTheorem} or \ref{SLCopt}; in particular, via biophysically plausible unsupervised sparse linear coding (e.g. \cite{rehnsommer2007, rozell2007neurally}).
\end{proposition}
%\cite{rehnsommer2007, rozell2007neurally, ganguli2012compressed, hu2014hebbian}

We close this section with a comment on sample complexity. A basic open problem remains: how many samples are necessary to determine the sparse linear coding model? If $k$ is held fixed or if the size of the support set of reconstructing codes is known to be polynomial in $\bar m$ and $k$, then a practical (polynomial) amount of data suffices.\footnote{In the latter case, a reexamination of the pigeonholing argument in the proof of Thm.~\ref{DeterministicUniquenessTheorem} requires a polynomial number of samples distributed over a polynomial number of supports.} Reasons to be skeptical that this holds in general, however, can be found in \cite{tillmann2014computational, Tillmann15}.

@@

In this section, we prove Lem.~\ref{MainLemma} after stating and proving some auxiliary lemmas.% and then sketch the proofs of Thm.~\ref{robustPolythm} and Cor.~\ref{ProbabilisticCor}.
